---
title: "Estimating Marginal Emissions of the Electrical Grid"
author: "Julian Ricardo"
date: "March 15, 2017"
output: html_document
runtime: shiny
bibliography: batt_lca_shiny.bib
csl: nature.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(plyr)
library(tidyverse)
library(lubridate)
library(reshape2)
asinh_trans <- function(){
  trans_new(name = 'asinh', transform = function(x) asinh(x), 
            inverse = function(x) sinh(x))
}
reverselog_trans <- function(base = exp(1)) {
  trans <- function(x) -log(x, base)
  inv <- function(x) base^(-x)
  trans_new(paste0("reverselog-", format(base)), trans, inv, 
            log_breaks(base = base), 
            domain = c(1e-100, Inf))
}
to_kg <- function(lb_val) {
  return(lb_val/2.205)
}
scrub_plants <- function(.data) {
  select(.data, 
         orispl, isorto, lat, lon,
         plprmfl, namepcap, plc2erta, capfac) %>% 
  mutate(namepcap = namepcap*0.87,
         plc2erta = to_kg(plc2erta),
         capfac = ifelse(capfac < 0, 0, capfac),
         fuel_type = case_when(
                                .$plprmfl == "WAT" ~ "Hydro",
                                .$plprmfl == "SUN" ~"PV",
                                .$plprmfl == "WND" ~ "Wind",
                                .$plprmfl == "NG" ~"Nat. Gas",
                                .$plprmfl == "NUC" ~ "Nuclear",
                                .$plprmfl == "LFG" ~ "Landfill Gas",
                                .$plprmfl %in% biomass ~ "Biomass",
                                .$plprmfl %in% petroleum ~ "Petro-fuels",
                                .$plprmfl %in% coal ~ "Coal-based",
                                .$plprmfl %in% other ~ "Other",
                                TRUE ~ .$plprmfl
                                )) %>% 
  filter(isorto != "",
         fuel_type != "")
}
make_dispatch <- function(.data) {
  mutate(.data,
         mc_rand = rnorm(nrow(.data), mc_m, mc_sd),
         mc_rand = ifelse(mc_rand < 0, 0, mc_rand)) %>%
    arrange(mc_rand) %>% 
    mutate(namepcap_cumul = cumsum(namepcap),
           plc2erta_wtd = namepcap*plc2erta,
           plc2erta_cumul = cumsum(plc2erta_wtd)/namepcap_cumul)
}
calc_plc2erta <- function(.data, dispatch) {
  runs <- unique(dispatch$run)

  cols <- cbind.data.frame(sapply(runs, function(n) {
           approx(x = filter(dispatch, run == n)$namepcap_cumul,
                  y = filter(dispatch, run == n)$plc2erta_cumul,
                  xout = .data$mw)$y})
  )
  names(cols) <- sapply(runs, function(n) paste0("plc2erta_", n))
  
  cbind.data.frame(.data, cols)
}
```

This R Markdown document will run through the initial procedure for estimating the time-dependent emissions impacts of using solar and energy storage to shave peak loads in buildings. The steps detailed below, using the NYISO[@NewYorkIndependentSystemOperator2014] electrical grid as a case study, involve:

1. Tidying power plant data
2. Adapting a cost-based model[@Gilmore2010] for determining dispatch order of power plants
3. Applying dispatch model to a grid load profile

Generally speaking, this method of estimating the emissions impacts of various induced changes in the consumption of grid electricity can be applied to:

- Renewable or distributed energy generation/storage
- Changes in energy consumption behavior
- Energy efficiency measures

Based on the work in this [repo](https://github.com/jric376/clca-batt), the focus here is on the first application. Following steps (e.g. setting up building energy consumption, solar, and storage models) are left for interactive notebooks to follow.

****

## Grid Inputs
```{r grid_input, include=FALSE}
coal <- c("BIT", "LIG", "SC", "SGC",
          "SUB", "WC")
petroleum <- c("BFG", "COG", "DFO", "JF",
               "KER", "PC", "RFO", "RG",
               "TDF", "WO")
biomass <- c("AB", "BLQ", "DG", "MSB",
             "MSW", "OBG", "OBS", "OG",
             "WDL", "WDS")
other <- c("GEO", "MWH", "PUR", "WH")
cbb_qual.n <- c("Biomass" = "#E69F00", "Coal-based" = "#999999",
                "Hydro" = "#CC79A7", "Landfill Gas" = "#009E73",
                "Nat. Gas" = "#F0E442", "Nuclear" = "#000000", 
                "Other" = "#762a83" , "Petro-fuels" = "#0072B2",
                "PV" = "#D55E00", "Wind" = "#56B4E9")
sim_results <- feather::read_feather("input/full_df.feather")
plant_costs <- data.table::fread("input/marg_costs.csv") %>% 
  select(-prim_mov,-c(heat_rt:VOM)) %>%
  transmute(MC, plprmfl = prim_fuel) %>% 
  group_by(plprmfl) %>%
  summarise(mc_m = mean(MC), mc_sd = sd(MC)) %>%
  mutate(mc_sd = case_when(is.na(.$mc_sd) ~ 0.05*.$mc_m,
                           TRUE ~ .$mc_sd))
theme_base <- theme(text = element_text(size = 14),
                    panel.background = element_rect(colour = "gray75", fill = "white")) +
                    theme(panel.grid.major = element_line(colour = "gray85")) +
                    theme(panel.grid.minor = element_line(colour = "gray85")) +
                    theme(legend.position = "none")
```

First, we load up data on power plants from the electrical grid of interest, based on territory name. We will focus on the nameplate capacities and CO~2~eq emissions rates of these plants, though the full dataset and accompanying documentation for the US EPA's eGrid2012[@eGRID2012] data are available [here](https://www.epa.gov/energy/emissions-generation-resource-integrated-database-egrid).

The code below also assigns marginal costs to each power plant by fuel type and generator type, calculated based on methods[@Gilmore2010] in the literature and publicly available data[@EIA2014b, @EIA2015, @EIA2014a]. In so doing, we estimate the priority with which power plants fire. Documented capacity factors may also be incorporated into future iterations of the grid dispatch model; they are displayed for reference only.

```{r scrub_plants}
plants <- data.table::fread("input/plants_all.csv") %>%
  scrub_plants() %>% 
  left_join(plant_costs, by = "plprmfl") %>% 
  filter(!is.na(mc_m))
```

```{r grid_plants, echo=FALSE}
df <- reactive({
  filter(plants,
         isorto == input$grid_terr,
         plants[[input$var]] > 0) %>% 
    group_by(fuel_type) %>% 
    mutate(Count = n()) %>% 
    select(fuel_type, Count,
           contains("namepcap"),
           contains("plc2erta"),
           contains("capfac")) %>% 
    summarise_if(is.numeric, c("median", "sd")) %>%
    mutate(Count_median = trunc(Count_median)) %>% 
    select(-Count_sd) %>% 
    rename("Fuel Type" = fuel_type,
           "Count" = Count_median,
           "Median MWh" = namepcap_median,
           "SD MWh" = namepcap_sd,
           "Median EF CO2eq" = plc2erta_median,
           "SD CO2eq" = plc2erta_sd,
           "Median Cap. Factor" = capfac_median,
           "SD Cap. Factor" = capfac_sd)
})

tabsetPanel(
  tabPanel("Graph",
          inputPanel(
                     selectInput("grid_terr", label = "Grid territory:",
                                 choices = unique(plants$isorto), selected = "NYISO"),
                      
                     radioButtons("var", "Variable to display:",
                                  c("Capacity (MW)" = "namepcap",
                                    "EF (kg CO2eq / MWh)" = "plc2erta",
                                    "Capacity Factor" = "capfac"))
                    ),

          boxplot <- renderPlot({ 
                      units <- switch(input$var,
                                      namepcap = "MW",
                                      plc2erta = "kg CO2eq / MWh",
                                      capfac = NULL)
                      
                      plants_terr <- filter(plants,
                                            isorto == input$grid_terr,
                                            plants[[input$var]] > 0)
                      
                      ggplot(plants_terr,
                             aes(x = factor(fuel_type),
                                 y = plants_terr[[input$var]],
                                 fill = factor(fuel_type))) +
                        geom_boxplot(varwidth = TRUE,
                                     outlier.shape = NA) +
                        geom_jitter(shape = 21, alpha = 1/2,
                                    position = position_jitter(w = 0.2)) +
                        scale_y_log10(limits = quantile(plants_terr[[input$var]], c(0.01, 0.99))) +
                        scale_fill_manual(name = "Fuel", values = cbb_qual.n) +
                        labs(x = NULL,
                             y = units) +
                        theme_base})
          ),
  tabPanel("Data", renderTable({df()}))
)
```

## NYISO case study

Depending on the degree of demand for electricity at the grid level, its operators (i.e. NYISO) allow bids for energy to be put forth until enough are submitted to satisfy the load. At this point, the grid operator pays all accepted bidders the highest marginal cost among all accepted bids.

Assuming that the marginal cost model approximates the subset of power plants that would actually fire at a given load level, we can then estimate the grid's cumulative emissions rate as well. To account for systematic uncertainties in the estimates, we compute emissions from multiple iterations of the dispatch model and display the resulting distribution. We use NYISO power plants for all following visualizations. 

### Drawing a dispatch order curve

```{r clean_dispatch}
nyiso_disp <- filter(plants, isorto == "NYISO") %>% 
  make_dispatch() %>% 
  mutate(run = 1)

disp_runs <- 10
full_disp <- nyiso_disp

for (i in 1:disp_runs) {
    disp <- make_dispatch(nyiso_disp) %>% 
      mutate(run = i)
    full_disp <- rbind.data.frame(full_disp, disp)
}
```

```{r dispatch, echo=FALSE, fig.height=10}


tabsetPanel(
  tabPanel("Graph",
          inputPanel(
            radioButtons("var2", "Variable to display:",
                         c("$ / kWh" = "mc_rand",
                           "Cumul. EF (g CO2eq / kWh)" = "plc2erta_cumul")),
            
            sliderInput("run_select", "Select a curve:",
                        1, 10, 1, ticks = FALSE)
            # actionButton("redraw", "Redraw curve",
            #   style="color: #fff; background-color: #999999; border-color: #999999")
          ),

          renderPlot({units <- switch(input$var2,
                                      mc_rand = "$ / kWh",
                                      plc2erta = "g CO2eq / kWh")
                      
                      input$run_select
                      
                      isolate ({
                        sample_run <- filter(full_disp, run == input$run_select)
                      })
                      
                      ggplot(sample_run,
                             aes(x = namepcap_cumul,
                                 y = sample_run[[input$var2]],
                                 fill = factor(fuel_type),
                                 size = namepcap)) +
                        geom_point(data = full_disp,
                                   aes(x = namepcap_cumul,
                                       y = full_disp[[input$var2]],
                                       fill = factor(fuel_type),
                                       size = namepcap),
                                   alpha = 1/3,
                                   color = "gray75") +
                        geom_point(shape = 21) +
                        scale_fill_manual(name = "Fuel",
                                          values = cbb_qual.n,
                                          guide = guide_legend(override.aes = list(size = 6))) +
                        scale_size(name = bquote(MW[plant]),
                                   breaks = c(100,500,1500),
                                   range = c(3,18)) +
                        labs(x = "",
                             y = units) +
                        theme_base +
                        theme(legend.position = "right",
                              legend.box = "vertical")
                      })
          ),
  tabPanel("Data",
           h3("\n"),
           output$downloadDispatch <- downloadHandler(
             filename = function() {
               "NYISO_plant_disp.csv"
             },
             content = function(con) {
              write.csv(disp, con) 
             }
           )
           )
)
```

### Visualizing Grid Load Profile

So far, we've only considered part of the problem. Knowing the subset of power plants that will fire is important, but the subset will change as costs and demands for energy fluctuate in time. We must therefore apply the dispatch curve model to a load profile[@NewYorkIndependentSystemOperator2014] of the electrical grid, as shown in the 'Weekly Load' tab below. From there, we can then generate time-varying estimates of the marginal emissions associated with grid electricity use throughout the year, as in the 'Quarterly Heatmap' tab.
```{r grid_load_setup, include = FALSE}
hr_labels <- unlist(lapply(seq(3,21,3),
                           function(x) ifelse(x>10,
                                              paste0(x, ":00"),
                                              paste0("0", x, ":00"))))
theme_heatmap <- theme(text = element_text(size = 14),
                       panel.background = element_blank(),
                       panel.border = element_blank(),
                       axis.line = element_blank(),
                       axis.ticks = element_blank(),
                       axis.text.y = element_text(angle = 33, hjust = 1),
                       axis.text.x = element_text(angle = 33, vjust = 1, hjust = 1),
                       strip.background = element_blank())
```

```{r grid_load_input}
nyiso_load.min <- data.table::fread("input/2014pal_combined.csv") %>% 
  transmute(mw,
            date_time = as.POSIXct(strptime(time_5min, format = "%m/%d/%Y %H:%M")),
            hr = hour(date_time),
            day = wday(date_time, label = TRUE),
            qtr = paste0("Q", quarter(date_time, with_year = FALSE)),
            wk = week(date_time)) %>% 
  calc_plc2erta(full_disp) %>% 
  gather(run, plc2erta, plc2erta_1:plc2erta_10)
```

```{r grid_load_plots, echo = FALSE}
units <- reactive({switch(input$load_param,
                          mw = "MW",
                          plc2erta = bquote(scriptstyle("g CO2eq / kWh")))})
         
nyiso_load.week <- reactive({
  nyiso_load.min %>%
   select(run, wk, date_time, mw, contains("plc2erta")) %>%
   filter(wk >= input$weeks[1] & wk <= input$weeks[2]) %>%
   mutate(mw_hr_avg = RcppRoll::roll_meanr(x = mw,
                                           n = 540,
                                           fill = "left"))
})

tabsetPanel(
  tabPanel("Weekly Load",
    inputPanel(
     sliderInput("weeks", "Weeks to display:",
                 1, 52, value = c(1,52))
    ),
    
    output$load_plot <- renderPlot({
     ggplot(nyiso_load.week(),
            aes(x = date_time,
                y = nyiso_load.week()[["mw"]])) +
         geom_line(size = 1.1) +
         geom_line(aes(y = nyiso_load.week()[["mw_hr_avg"]]),
                   colour = cbb_qual.n[["PV"]],
                   size = 1.33) +
         labs(x = "",
              y = "MW") +
         theme_base
    })
  ),
  tabPanel("Quarterly Heatmap",
           inputPanel(
             radioButtons("load_param", "Quantity to display:",
                           c("MW" = "mw",
                             "EF (kg CO2eq / MWh)" = "plc2erta")),
             radioButtons("var3", "Mean or Std. Dev.:",
                          c("Mean" = "mean",
                            "Std. Dev." = "sd"))
           ),
           
           output$heatmap <- renderPlot({
            nyiso_load.var <- nyiso_load.min %>%
                group_by(run, qtr, day, hr) %>% 
                select(run, qtr, day, hr,
                       -date_time, -wk,
                       contains(input$load_param)) %>% 
                summarise_all(c("mean", "sd")) %>%
                select(run, qtr, day, hr, contains(input$var3)) %>% 
                melt(id.vars = c("qtr", "day", "hr"),
                     measure.vars = input$var3) %>%
                ddply(.(variable), transform, rescale = scale(value)) %>% 
                filter(variable == input$var3,
                       !is.na(day))
             
             ggplot(nyiso_load.var,
                    aes(y = day, x = hr, fill = value)) +
               facet_wrap("qtr") +
               geom_tile(color = "gray85") +
               scale_x_continuous(breaks = seq(2,20,3),
                                  labels = hr_labels,
                                  expand = c(0,0)) +
               scale_fill_gradient2(name = units(), low = "#7b3294", high = "#008837",
                                    midpoint = mean(nyiso_load.var$value)) +
               labs(x = "",
                    y = "") +
               theme_heatmap
           })
  )
)
```


## Future work

- Accounting to some extent for geospatial constraints of electrical grid
- Address apparent lack of petro-fuels plants firing
- Incorporating published capacity factors into with marginal cost model
- Compile load profiles from different electric grids and years

****

## References

