---
title: "Peak shaving power flows"
author: "Julian Ricardo"
date: "May 7, 2017"
output: html_document
runtime: shiny
bibliography: batt_lca_shiny.bib
csl: nature.csl
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dygraphs); library(xts); library(zoo)
library(plyr)
library(tidyverse)
library(lubridate)
library(reshape2)
library(xts)
market_ts.raw <- list.files("input", pattern = "market.*run\\.csv", full.names = TRUE) %>% 
  read.csv() %>% 
  mutate(X = NULL,
         date_time = ymd_hm(date_time))

dmd_thresh <- max(market_ts.raw$bldg_kw)*0.5
rand_seed <- 12387
```

# Sample simulation results

Every building simulation generates time-series representing the building load profile before and after peak shaving, in addition to the generation and (dis)charge profiles for the solar and battery systems, respectively. We can visualize these concurrent time-series using a sample run from a simulated supermarket, shaving peak demand by $50\%$ in conjunction with a 487 kWh vanadium redox flow battery and a 234 kWh photovoltaic array.

### Load profiles

In the figures below, the red dotted line represents the peak shaving threshold for the sample time-series, set at $50\%$ of annual peak demand: `r round(dmd_thresh, 0)` kW.

```{r load_ts_prep, echo=FALSE}
set.seed(rand_seed)
winter_sample <- market_ts.raw %>% 
  filter(month(date_time) == 1) %>% 
  mutate(wk = week(date_time)) %>% 
  filter(wk == sample(unique(.$wk), 1)) %>% 
  select(-wk, -curtail_kw)

summer_sample <- market_ts.raw %>% 
  filter(month(date_time) == 8) %>% 
  mutate( wk = week(date_time)) %>% 
  filter(wk == sample(unique(.$wk), 1)) %>% 
  select(-wk, -curtail_kw)

winter_xts <- winter_sample %>% 
  select(date_time, ends_with("_kw")) %>% 
  tidyquant::as_xts(date_time)
winter_soc <- winter_sample %>% 
  select(date_time, soc) %>%
  tidyquant::as_xts(date_time)

summer_xts <- summer_sample %>% 
  select(date_time, ends_with("_kw")) %>%
  tidyquant::as_xts(date_time)
summer_soc <- summer_sample %>% 
  select(date_time, soc) %>%
  tidyquant::as_xts(date_time)
```


```{r load_ts, echo=FALSE}
tabsetPanel(
  tabPanel("Winter",
            output$winter_xts <-  renderDygraph({
              dygraph(winter_xts) %>% 
                dyAxis("y", label = "kW") %>%
                dyLimit(dmd_thresh, color = "red") %>% 
                dyLegend(width = 300) %>% 
                dyRoller(rollPeriod = 12) %>% 
                dyOptions(drawGrid = FALSE, fillGraph = TRUE,
                         colors = RColorBrewer::brewer.pal(5, "Dark2"))
            }),
            output$winter_soc <- renderDygraph({
              dygraph(winter_soc) %>% 
                dySeries("V1", label = "SoC") %>% 
                dyAxis("y", label = "SoC") %>%
                dyRoller(rollPeriod = 12) %>% 
                dyOptions(drawGrid = FALSE, fillGraph = TRUE,
                         colors = RColorBrewer::brewer.pal(1, "Dark2"))
            })
  ),
  tabPanel("Summer",
           output$summer_xts <- renderDygraph({
             dygraph(summer_xts) %>% 
               dyAxis("y", label = "kW") %>%
               dyLimit(dmd_thresh, color = "red") %>% 
               dyLegend(width = 300) %>% 
                dyRoller(rollPeriod = 12) %>% 
               dyOptions(drawGrid = FALSE, fillGraph = TRUE,
                         colors = RColorBrewer::brewer.pal(5, "Dark2"))
           }),
           output$summer_soc <- renderDygraph({
             dygraph(summer_soc) %>% 
               dySeries("V1", label = "SoC") %>% 
               dyAxis("y", label = "SoC") %>%
               dyRoller(rollPeriod = 12) %>% 
               dyOptions(drawGrid = FALSE, fillGraph = TRUE,
                         colors = RColorBrewer::brewer.pal(1, "Dark2"))
           })
  )
)
```

# Summary
